Predicting Success of Motivational Interviews with Recurrent Neural Network and Probabilistic Model for Patient-Provider Communication Sequences

### 2017-09-11 paper-amia-summit-2018/amia_summit_2018.tex
update discussion section

\section*{Discussion}
By analyzing the experimental results of different communication sequence outcome prediction schemes proposed in this paper, we arrived at the following conclusions. First, the overall predictive performance of RNN models is substantially better than probabilistic models. In particular, RNN-based method achieves near-human accuracy for predicting the success of motivational interviews. This indicates that RNN are able to capture the structure of discourse in motivational interviews by preserving long-term dependencies among the behavior codes, which reflect the overall progression of the interviews. In addition to that word embedding allows to reduce the dimensionality of codes in PPC sequences and consequently improve both precision and recall of the prediction method.

Second, computing loss with target replication at each time step results in better performance for all configurations of the proposed RNN method. This indicates that mean of the losses over all steps emphasize the dependencies between the pairs of patient and provider codes, which results in more accurate estimates of the model parameters. Better estimates of parameters in RNN models of motivational interviews are propagated to the next step based on the relative importance of intermediate output, where they are aggregated into predictions for the entire sequence. This allows to achieve an improvement in the prediction accuracy of the method.

@@

Fourth, similar to traditional markov model, HMM achieve a dramatic improvement in the prediction accuracy when larger training set is used. This indicates that sufficient training data is required to estimate optimal hyperparameters such as number of hidden states, intitial state distribution, transition probabilities, and emission probabilities.

Fifth, the proposed method can be used to identify the most effective communication strategies at eliciting a particular type of behavioral response. Awareness of these strategies by researchers can significantly decrease the time and effort required to develop effective interventions to address many public health conditions, such as childhood obesity, and tailor these interventions to particular patient cohorts. Awareness of these strategies by the counselors can lead to greater success rate of motivational interviews.

>>>>

\section*{Discussion}
By analyzing the experimental results of different communication sequence outcome prediction schemes proposed in this paper, we arrived at the following conclusions. First, the overall predictive performance of RNN models is substantially better than probabilistic models. In particular, the RNN-based method achieves near-human accuracy for predicting the success of motivational interviews. This indicates that RNN is able to capture the structure of discourse in motivational interviews by preserving long-term dependencies among the behavior codes, which reflect the overall progression of the interviews. In addition to that word-embedding allows to reduce the dimensionality of codes in PPC sequences and consequently improve both precision and recall of the prediction method.

Second, computing loss with target replication at each time step results in better performance for all configurations of the proposed RNN method. This indicates that mean of the losses over all steps emphasize the dependencies between the pairs of patient and provider codes, which results in more accurate estimates of the model parameters. Better estimates of parameters in RNN models of motivational interviews are propagated to the next step based on the relative importance of intermediate output, where they are aggregated into predictions for the entire sequence. This allows achieving an improvement in the prediction accuracy of the method.

@@

Fourth, similar to traditional Markov model, HMM achieve a dramatic improvement in the prediction accuracy when larger training set is used. This indicates that sufficient training data is required to estimate optimal hyper parameters such as a number of hidden states, initial state distribution, transition probabilities, and emission probabilities.

Fifth, the proposed method can be used to identify the most effective communication strategies at eliciting a particular type of behavioral response. Awareness of these strategies by researchers can significantly decrease the time and effort required to develop effective interventions to address many public health conditions, such as childhood obesity, and tailor these interventions to particular patient cohorts. Awareness of these strategies by the counselors can lead to a greater success rate of motivational interviews.
